---
title: "ST558 HW6"
author: "Rocco Matarazzo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(parallel)
```
# Our Very Own T-Test

First, we'll write a function that takes on a vector of numeric values and the mean value of comparsion that ultimately outputs the observed t-statistic.

```{r}

t_observed <- function(data, mu_0){
  # Getting the Sample Mean
  sampleMean <- mean(data)
  # Getting the sample std. dev.
  sampleSD <- sd(data)
  # Getting the sample size
  n <- length(data)
  
  # Getting the observed t-Stat
  t_obs <- (sampleMean - mu_0)/(sampleSD/sqrt(n))
  # Printing the observed stat.
  return(t_obs) 
}

```

Our next function will tell us whether or not to reject the null hypothesis. This function will take the inputs of the observed test statistic, sample size, significance level (alpha), and direction of the test.

```{r}

decision <- function(t_obs, n, alpha, direction){
  # Get Critical Value for Rejection
  if(direction == "left" | direction == "Left"){
    
    # HA: less than symbol
    critValue <- qt(alpha, n-1, lower.tail = FALSE)
   # Compare this with t_obs
    if(t_obs < critValue){
      result <- TRUE
    }else{
      result <- FALSE
    }
  }
  else if(direction == "right" | direction == "Right"){
        
    # HA: greater than symbol
    critValue <- qt(1-alpha, n-1, lower.tail = FALSE)
   # Compare this with t_obs
    if(t_obs > critValue){
      result <- TRUE
    }else{
      result <- FALSE
    }
  }
  else if(direction == "two-sided" | direction == "Two-Sided" | direction == "Two Sided"){
        
    # HA: not equal than symbol
    critValue <- qt(1-(alpha/2), n-1, lower.tail = FALSE)
    # Compare this with t_obs
    if(abs(t_obs) > critValue){
      result <- TRUE
    }else{
      result <- FALSE
    }
  }else{
  # if none options available   
    break
  
  }
  
  return(result)
}

```

Now we can test out our two functions with the classic iris dataset.

### Sepal Length != 5.5
```{r}

# Getting the observed t-stat
t_obs <- t_observed(iris$Sepal.Length, 5.5)

# Getting decision to reject or not
decision(t_obs, length(iris$Sepal.Length), 0.05, direction = "two-sided")
```
Here we shall reject the null hypothesis. There is evidence to suggest that the mean sepal length differs from 5.5.


### Sepal Width > 3.5
```{r}

# Getting the observed t-stat
t_obs <- t_observed(iris$Sepal.Width, 3.5)

# Getting decision to reject or not
decision(t_obs, length(iris$Sepal.Width), 0.05, direction = "Right")
```
Here we shall fail to reject the null hypothesis. There is not evidence to suggest that the mean sepal width is greater than 3.5.


### Petal Length < 4
```{r}

# Getting the observed t-stat
t_obs <- t_observed(iris$Petal.Length, 3.5)

# Getting decision to reject or not
decision(t_obs, length(iris$Petal.Length), 0.05, direction = "Left")
```
Here we shall fail to reject the null hypothesis. There is evidence to suggest that the mean petal length differs from 5.5.

# Monte Carlo Study

First, we shall generate many random values from the Gamma Distribution.
```{r}
# Setting Parameters
n <- 100
shape <- 20
rate <- 1

# number of sims
nsim <- 10000

# creating a function that we can replicate over and over again
testResults <- function(n, shape, rate){
    null_mean <- shape*rate
    randomValues <- rgamma(n, shape, rate)
    testStat <- t_observed(randomValues, null_mean)
    outcome <- decision(testStat, n, 0.05, "Two-Sided")
    outcome
}

# Replication taking place (Turning this into a function):
replicationFunction <- function(list_object) {
  rate <- 1 # keeping rate constant at 1 
  nsim <- 10000 #number of simulations
  
  # creating blank DF to store output
  output <- data.frame(matrix(ncol = 3, nrow = 0))
  x <- c("n", "shape", "mean_proportion")
  colnames(output) <- x
  
  for(i in 1:length(list_object)){
    
  n <- list_object[[i]]$n 
  shape <- list_object[[i]]$shape
  replication <- replicate(nsim, testResults(n, shape, rate))

  # output[i,]$n <- n
  # output[i,]$shape <- shape
  # output[i,]$mean_proportion <- mean(replication)
  # # Getting the means
  }  
  return(mean(replication))
    }
```

# Parallel Computing
```{r, error = TRUE}
# Setting Base Vectors
sampleSizes <- c(10, 20, 30, 50, 100)
shapes <- c(0.2, 0.5, 1, 2, 5, 10, 20)
rates <- c(1)

# Getting a dataframe with each combination
df <- expand.grid(n = sampleSizes, shape = shapes)

# Using the purrr package
# I believe this is possible as Dr. Post suggests with apply() and such
# but I have used this solution before
passingValues <- df %>% purrr::transpose()

# Making my cluster
cores <- detectCores() # ive got 8 cores! 
cluster <- makeCluster(cores - 1) # setting up 7 cores
clusterExport(cluster, list("replicationFunction", "testResults", "t_observed", "decision")) # setting up functions

# Doing the function through parallelization
resultsPar <- parLapply(cluster, 
                      X = passingValues,
                      fun =  replicationFunction(list_object = passingValues)
                      )
```
Unfortunately I am still getting an error here. Due to time constraints this week in particular (exam in other class and more than usual work appointments) I was unable to complete this. I appreciate your understanding and understand the potential for a low grade for this assignment.